MASU Paper

### 2022-05-30 paper/paper_draft.Rmd
updated introduction and added an abstract

@@

Due to the history of hydrologic modifications such as the development of reservoirs, dams, and irrigation systems, there has been an apparent decrease in the number of migratory fishes in the recent years. For example, on the Pacific coast there has been significant decreases in diversity within salmonid species due hydrologic and dam alterations. Atlantic salmon and anadromous species of sturgeons are also impacted negatively because of a loss of portions to their habitats upstream of dams. Overall, the diversity within migratory fish species have largely declined due to these dams blocking the fishes' migratory paths. One species of migratory fish that we will be discussing more in depth is the MASU salmon.

MASU salmon, or the cherry salmon, have the most limited distribution among the Pacific salmon. Additionally, since they spend a lot of their life in freshwater rivers or streams, this makes them more inclined to be exposed to fishing and sportfishing pressures. Since they are more vulnerable than other salmon species, it is important to look into the details that influence their life cycles of being either resident or anadromous. Anadromous salmon are initially born in the rivers then migrate to the seas where they develop larger body sizes. These salmon eventually migrate back to the rivers in order to breed. The resident salmon in contrast sexually mature at a younger age and stay within the river systems throughout their entire lives (Morita and Nagasawa)

MASU salmon and their life cycles have been known to be primarily influenced by environmental factors such as temperature and other factors such as the salmon's body size, growth rate and lipid stores. Environmental factors such as climate change have been known to impact the decision of the salmon to be either resident or anadromous. It has been shown that the number of resident males have increased while the proportion of anadromous males and the proportion of delayed migrants within both females and males have decreased as the temperatures have increased. The increase in resident males due to increasing temperatures have led to a significant decrease in the number of anadromous individuals. Instead of looking into environmental influences, which have already beeen heavily studied, in this paper, we will be examining if there is a genetic basis within MASU salmon that will allow us to determine the salmon's life cycle.

For comparison, within the O. mykiss species or rainbow trout, which is also a species of fish that exhibit either resident or anadromous life cycles, there is strong evidence suggesting a genetic basis on life history. The first evidence of a genetic basis for O. mykiss was when Neave (1944) conducted common garden and transplant experiments with sympatric rainbow trout and steelhead. After marking and releasing the progeny of the rainbow trout and steelhead, he found that life cycles were heritable after a larger proportion of rainbow trout offspring were recaptured over many years. Additionally, crosses between anadromous O. mykiss individuals resulted in the most smolts while crosses between resident individuals produced the fewest number of smolts. In a recent study by Hess et al. (2016), univariate analyses and a multivariate random forest (RF) machine learning algorithm was used to perform the association mapping of of 15 239 single nucleotide (SNPs) for adult migration-timing phenotype in steelhead. Using a univariate linear model resulted in the finding of three SNPs from a gene that was strongly linked with migration timing. Using multivariate analyses with RF resulted in the finding of 15 additional SNPs that were connected to 60% of the trait variation.

Like how the rainbow trout and steelhead have a clear genetic basis for their life histories, we also wanted to investigate the genetic influences on MASU salmon and their life cycles using similar methods. We performed a random forest on the data, did association tests, and found diversity statistics, which did show any clear indicator of the genetic difference between the anadromous and resident MASU salmon. However, when creating a Principle Component Analysis chart (PCA) to look for patterns in relatedness between resident and anadromous individuals, we found that the anadromous individuals clustered significantly closer to the cluster center for individuals from the other population. Overall, from the DNA analysis methods used in this paper, there is not strong evidence that there is a genetic basis for MASU salmon life histories, but further studies using a bigger sample size would need to be conducted in order to investigate the influences of genetics on MASU salmon and their life cycles.

>>>>

#Abstract
Within Masu salmon (Oncorhynchus masou), individuals can exhibit one of two life cycles: resident or anadromous. The expresson of either life history is mainly influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced (CITE). Here, we explore if there are genetic factors that influence the development of either residency or anadromy. In order to look for patterns of relatedness among resident and anadromous individuals, we made a PCA chart, did association tests, and made a random forest (RF). Ultimately, we did not find any significant associated genetic elements, but further research with a larger sample size is needed in order to fully assess potential genetic influences on anadromy and residency.

@@

Due to the history of hydrologic modifications such as the development of reservoirs, dams, and irrigation systems, there has been an apparent decrease in the number of migratory fishes in the recent years (CITE). In North America, for example,there have been significant decreases indiversity within pacific salmonid species due hydrologic and dam alterations (CITE), and atlantic salmon and anadromous species of sturgeons are also impacted negatively because of a loss of portions to their habitats upstream of dams (CITE). Overall, the diversity within migratory fish species have largely declined by an average of 70% between 1970 and 2016 due to these dams blocking the fishes' migratory paths (CITE- need a research paper for this number)

Masu salmon (Oncorhynchus masou, otherwise known as cherry salmon) have the most limited distribution among the Pacific salmon (CITE). Since they spend a greater portion of their life in freshwater rivers or streams than most other migratory salmon, they are particularly more exposed to fishing and sportfishing pressures (CITE).Their increased vulnerability makes it critical to understand the details that influence and control migratory life history variance in the species. Masu in particular (like Oncorhynchus mykiss, steelhead salmon/rainbow trout) are partially migratory, with large resident and anadromous populations (CITE). As is typical, anadromous masu are initially born in the rivers then migrate to the seas where they develop larger body sizes, then eventually migrate back to their natal rivers in order to breed (CITE). Resident masu in contrast sexually mature at a younger age and stay within the river systems throughout their entire lives (Morita and Nagasawa)

The decision to migrate in masu salmon is thought to be primarily influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced (CITE). Consequentially, changes in environmental factors, such as those caused by climate change, have been known to impact the decision of the masu to be either resident or anadromous (CITE). For example, X et al (YEAR) has shown that the number of resident males have increased over the last X years while the proportion of anadromous males and the proportion of delayed migrants within both females and males have decreased as the temperatures have increased, which has led to a significant decrease in the number of anadromous individuals overall.

In contrast, the genetic factors that contribute to migration in masu salmon have not been well studied.  Despite the prevalence of environmental control of migration in masu, it is not unreasonable to expect genetic elements to play some role in governing migratory decisions in the species. For example, within steelhead, there is strong evidence suggesting both an environmental and genetic basis for migratory life history. In 1944, Naeve conducted common garden and transplant experiments with sympatric rainbow trout and steelhead and found that migratory status was heritable after observing that rainbow trout offspring were considerably more likely to be resident themselves, whereas  crosses between anadromous individuals resulted in the most anadromous smolts. It is not surprising, therefore, that a recent study by Hess et al. (2016), on 15,239 single nucleotide polymorphisms (SNPs) found of three SNPs from a gene that was strongly linked with adult migration timing, and that 18 total SNPs that were connected to 60% of the trait variation. Despite this, however, growth rates are known to strongly influence anadromy in the species—individuals which grow faster as juveniles are far more likely to stay as residents than otherwise (CITE). Indeed, migration in the species seems to be controlled jointly the interplay of genetic and environmental factors:

The degree of genetic control of migration in masu, however, is still not well known. To improve our understanding of the mechanisms behind residency and anadromy in the species,  we sequenced a set of anadromous and resident masu individuals using high-resolution Restriction Associated Digest (RAD) sequencing.  However, we did not find any clear indicator of the genetic difference between ouranadromous and resident Masu salmon. While this may suggest that there is not that there is a strong genetic basis for Masu salmon migratory life history variation, further studies using a bigger sample size from a larger geographic range would need to be conducted in order to fully investigate the influences of genetics on Masu salmon migratory life cycles.
